Timeliness of security updates for Android

### 2015-03-18 securityupdates.tex
Attempt to make things work better with the single column format

\label{sec:android_update_process}
\begin{figure}
 \centering
 \def\svgwidth{\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, numbers in brackets represent number of such entities in our data.
 Dotted arrows indicate flows where we do not know how many updates are being produced as we can't measure those flows directly as they are not public.\todolater{Turn into a sankey diagram}}
 \label{fig:update_ecosystem}
\end{figure}
To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
Consequently Android can include any compatible versions of those projects, including those which fix security vulnerabilities.
Android also incorporates various drivers for different bits of hardware.
The Android platform is then built from these components by Google.
The code for each Android release or update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}} until after a binary release has been published.
Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
The network operator may then make or request further customisations and perform further testing before shipping the update to the Device.
Sometimes device manufactures ship updates directly to the user, sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices.
Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.

@@

\begin{figure}
 \includegraphics[width=\columnwidth]{figures/from_to_updates.pdf}
 \caption{Updates between different Android versions in the \da\ data}
 \label{fig:from_to_updates}
\end{figure}
\begin{figure}
 \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
 \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
 \label{fig:weekly_security_updates}
\end{figure}
We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngradesNominal) when older versions of Android are installed on to devices.
Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.

>>>>

\label{sec:android_update_process}
\begin{figure}[h]
 \centering
 \def\svgwidth{0.7\columnwidth}
 \import{figures/}{update_ecosystem.pdf_tex}
 \caption{Flow of updates between participants in the Android ecosystem.
 Numbers on edges indicate updates shipped between \daStartDate\ and \daEndDate, those in brackets represent number of such entities.
 Dotted arrows indicate flows where we can't measure how many updates are produced as they are not public.}
 \label{fig:update_ecosystem}
\end{figure}
To understand how vulnerabilities in Android are fixed we must examine the Android update process which we model in Figure~\ref{fig:update_ecosystem}.
There are five entities or groups which contribute towards Android updates: the network operators, the device manufacturers, the hardware developers, Google and the upstream open source projects.
Android builds on various open source projects such as the Linux kernel, OpenSSL and BouncyCastle cryptography libraries.
Consequently Android can include any compatible versions of those projects, including those which fix security vulnerabilities.
Android also incorporates various drivers for different bits of hardware.
The Android platform is then built from these components by Google.
The code for each Android release or update is kept secret\footnote{\url{https://source.android.com/source/code-lines.html}} until after a binary release has been published.
Device manufacturers likely receive advanced access in order to prepare handsets so they can customise it before passing it on to the network operator.
The network operator may then make or request further customisations and perform further testing before shipping the update to the Device.
Sometimes device manufactures ship updates directly to the user, sometimes the device manufacturer and Google collaborate closely to make a particular phone, such as with Nexus devices.
Sometimes device manufacturers incorporate upstream open source project releases directly, and sometimes incorrectly -- for example previous work has recorded evidence of broken nightly builds of sqlite in Android releases~\cite{Wagner2013}.

@@

\begin{figure}
 \centering
 \includegraphics[width=0.6\columnwidth]{figures/from_to_updates.pdf}
 \caption{Updates between different Android versions in the \da\ data}
 \label{fig:from_to_updates}
\end{figure}
\begin{figure}[h]
 \includegraphics[width=\columnwidth]{figures/nw_security_updates.pdf}
 \caption{Proportion of updates each week which fixed or may have fixed security vulnerabilities}
 \label{fig:weekly_security_updates}
\end{figure}
We recorded update events and Figure~\ref{fig:from_to_updates} shows how devices upgrade between different versions of Android.
Mostly the dark cells are upgrades (above the diagonal) (\daNumUpdatesUpgrades).
While many upgrades are from one version to the next version there are also a fair number (\daNumUpdatesBigUpgrades, \daPercBigUpgradesNominal) which skip more than \daNumUpdatesSkippedBig\ versions.
Surprisingly there are also a small number of downgrade events (\daNumUpdatesDowngrades, \daPercUpdatesDowngradesNominal) when older versions of Android are installed on to devices.
Possible reasons why users are downgrading are to free up space on their device, to make it easier to root or because a new version introduced bugs.
